---
title: "King Auric's Options"
date: "6/20/2020"
author: "Aaron Ferrucci"
output:
  html_document:
    df_print: paged
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r message=F, warning=F, echo=F}
library(dplyr)
library(usmap)
library(ggplot2)
```
From [538 Riddler: Can You Flip The Magic Coin?](https://fivethirtyeight.com/features/can-you-flip-the-magic-coin/)

> From Dean Ballard comes a riddle of radiant spheres and fatherhood,
> just in time for the summer solstice and Father’s Day:
> 
> King Auric adored his most prized possession: a set of perfect spheres of
> solid gold. There was one of each size, with diameters of 1 centimeter, 2
> centimeters, 3 centimeters, and so on. Their brilliant beauty brought joy
> to his heart. After many years, he felt the time had finally come to pass
> the golden spheres down to the next generation — his three children.
> 
> He decided it was best to give each child precisely one-third of the
> total gold by weight, but he had a difficult time determining just how
> to do that. After some trial and error, he managed to divide his spheres
> into three groups of equal weight. He was further amused when he realized
> that his collection contained the minimum number of spheres needed for
> this division. How many golden spheres did King Auric have?
> 
> Extra credit: How many spheres would the king have needed to be able to
> divide his collection among other numbers of children: two, four, five,
> six or even more?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

