---
title: "King Auric's Options"
date: "6/20/2020"
author: "Aaron Ferrucci"
output:
  html_document:
    df_print: paged
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r message=F, warning=F, echo=F}
library(dplyr)
library(ggplot2)
library(utils)
```
From [538 Riddler: Can You Flip The Magic Coin?](https://fivethirtyeight.com/features/can-you-flip-the-magic-coin/)

> From Dean Ballard comes a riddle of radiant spheres and fatherhood,
> just in time for the summer solstice and Father’s Day:
>
> King Auric adored his most prized possession: a set of perfect spheres of
> solid gold. There was one of each size, with diameters of 1 centimeter, 2
> centimeters, 3 centimeters, and so on. Their brilliant beauty brought joy
> to his heart. After many years, he felt the time had finally come to pass
> the golden spheres down to the next generation — his three children.
>
> He decided it was best to give each child precisely one-third of the
> total gold by weight, but he had a difficult time determining just how
> to do that. After some trial and error, he managed to divide his spheres
> into three groups of equal weight. He was further amused when he realized
> that his collection contained the minimum number of spheres needed for
> this division. How many golden spheres did King Auric have?
>
> Extra credit: How many spheres would the king have needed to be able to
> divide his collection among other numbers of children: two, four, five,
> six or even more?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Preliminaries
Thinking about the problem, what do we know?

* We need at least 3 balls (more generally, at least one ball per child)
* The weight of the gold balls is proportionate to the cube of the diameter. (I can work with cubed diameters (d3s) rather than actually ball weights.)
* The sum of the d3s must be a multiple of 3

### Nuts 'n' Bolts
How common is it to find a sum of d3s that's divisible by 3?
```{r}
n <- 30
d3s <- seq(n)**3
sums <- cumsum(d3s)
sums %% 3
```

Looks like 2/3s of the sums are divisible by 3. So, a small optimization here: skip the 1/3 of cases which fail the divisible-by-3 test.


A solution to this problem (a brute force one, anyway) could rely on finding a subset of a list of numbers which sums to a specific value. Really brute force would be to compute each subset, do the sum, and compare. There's a better way via dynamic programming, I'm sure.

```{r}
# find all the combinations of the input list of a particular size that sum to a given value
#   nums: numeric list of values for forming the sums
#   count: number of elements from for the candidate sums
#   required_sum: what the combination(s) must add up to
get_summing_combs <- function(nums, count, required_sum) {
  # default return value is NULL (meaning no combinations were found)
  summing_combs <- NULL
  # Get all the combinations of the 'nums' array that have 'count' elements
  # this is a matrix with columns representing combinations
  combinations <- combn(nums, count)
  # add up all the columns
  all_column_sums <- apply(combinations, 2, sum)
  # check all those sums: do any match the requirement?
  if (any(all_column_sums == required_sum)) {
    # found at least one match, save it or them
    # force single-match vector into Nx1 matrix; otherwise the result is already a matrix, and as.matrix is a no-op.
    summing_combs <- as.matrix(combinations[,all_column_sums == required_sum])
  }
  return(summing_combs)
}

# given a list of values proportionate to ball volume, find the smallest number of balls that
# can be partitioned into 3 equal-volume lists.
# The return value is a list of <ball_count>, then 3 vectors for the partitions.
# If none are found, return NULL
find_shares <- function(d3s) {
  # Start at count of 3 - no way to split 1 or 2 balls into 3 equal partitions.
  for (i in seq(3, length(d3s))) {
    # minor optimization: skip if the total is not a multiple of 3
    if (sums[i] %% 3 == 0) {
      find_sum <- sums[i] / 3 # the sought sum
      cur_d3s <- d3s[1:i] # the list of i balls to consider in this step
      # now consider all the combinations of the balls, from size 1 up to <num balls> - 2
      # (no need to consider those largest combinations, because they wouldn't leave any balls
      # to be shared by the other two heirs).
      for (j in seq(i - 2)) {
        # get a matrix of all the combinations of size j that sum to the required value.
        match1 <- get_summing_combs(cur_d3s, j, find_sum)
        if (!is.null(match1)) {
          # found one or more combinations
          # next try to find combinations in the complement set - the remaining balls
          for (k in seq(ncol(match1))) {
            compl <- cur_d3s[!cur_d3s %in% match1[,k]]
            # look at all combinations of the complement set from 1 up to <size - 1>
            # (like above, don't consider the largest combination, because that would leave
            # nothing for the remaining heir)
            for (l in seq(length(compl) - 1)) {
              match2 <- get_summing_combs(compl, l, find_sum)
              if (!is.null(match2)) {
                # found at least one match in the complement set. The remainder must
                # also be a sum, so this is a solution. Let's choose the first one.
                # (It's possible that any remaining alternates are equivalent)
                # bundle stuff up to be returned
                share1 <- match1[,k]
                share2 <- match2[,1]
                share3 <- compl[!compl %in% share2]

                return(list(ball_count=i, share1=share1, share2=share2, share3=share3))
              }
            }
          }
        }
      }
    }
  }
  return(NULL)
}
solution <- find_shares(d3s)
if (is.null(solution)) {
  print("sorry, no solution was found")
} else {
  print(solution)
}
```
King Auric has `r solution$ball_count` gold balls, with diameter 1cm, 2cm, ..., `r solution$ball_count`cm. And he can divide them equally among his three children! Some sanity checking:
```{r}
sums[solution$ball_count] / 3
sum(solution$share1)
sum(solution$share2)
sum(solution$share3)
```
Looks like all three shares add up to the required amount. But is there any overlap?
```{r}
intersect(solution$share1, solution$share2)
intersect(solution$share1, solution$share3)
intersect(solution$share2, solution$share3)
```

One more test, does all the shares together include all the balls?
```{r}
un <- union(union(solution$share1, solution$share2), union(solution$share1, solution$share3))
un[sort.list(un)]
all(un[sort.list(un)] == d3s[1:solution$ball_count])
```

That's that, seems like _a_ solution. I haven't shown that it's a solution with the minimal number of balls; if there's a bug in my code, I could have missed a smaller solution.

